---
title: 'Homework #3'
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A useful note
Remove any `str_view()`, `View()` or `install.packages()` commands in your `Rmd` file. These commands will prevent the knitting of your file. 

## Strings
1.  Create a variable storing single string containing the following paragraph exactly including formatting. Display the variable in a reader-friendly manner.
    
      ```
      "Age is an issue of mind over matter. If you don't mind, it doesn't matter."
      
      -- Mark Twain
      ```

2. Write code to tell whether each entry in the following string has identical first and second halves. Do not consider cases here, so both `abcdabcd` and `abcdABCD` have identical first and second halves.

    ```
    c('apple', 'appleapple', 'BananaBanana', 'Banana', 'cherryCherry')
    ```

3. Count the number of words in the following string. [Hint: You can use `str_split` with a well-designed pattern to separate the words. `str_split(x, pattern)[[1]]` extracts the vector of words.]

    ```
    "Dear Friends,

    The exuberance of the change of seasons at Iowa State University never grows old — from the blooms of spring to the joys of another graduating class. During this holiday season, let your spirits be lifted by family and friends — and by this short video of captivating images of this magnificent campus.
    
    Our best wishes to you and yours for a happy, hope-filled 2022!

    Wendy Wintersteen and Robert Waggoner"

    ```

4. You will be analyzing the movie script of The Matrix (1999) in this question. The script can be found [here](....). Download the file and put it side by side your source file. Then use `script <- readLines('matrix_script.txt')` to read in the data. [Hint: If the command does not work, you may have a wrong working directory. In RStudio, click Session -- Set Working Directory -- To Source File Location]
    
    1. Look at the script file and see what it has. Compare it with the `script` variable. This part is for your own understanding and you don't need to answer anything.
    
    2. How many times did Neo speak?
    
    ```{r, include=FALSE, eval=FALSE}
    library(stringr)
    script <- readLines('matrix_script.txt')
    neo <- str_detect(script, '^NEO')
    table(neo)
    ```
    
    3. [*Optional for STAT 486 students*] How many time did Neo refer to Trinity?
    
    
## Factors and Logical variables

TODO: post dataset.
You will be working with an extraction related to happiness from the general social survey (GSS) available [here](happy.csv). The GSS is a yearly survey of Americans. The data you will work on spans 1972 to 2018.
Use `happy <- readr::read_csv('happy.csv')` to read in the data.

```{r, include=FALSE, eval=FALSE}
library(dplyr)
write.csv(happy %>% select(-polviews), file='happy.csv', row.names = FALSE)
happy <- readr::read_csv('happy.csv')
```

1. Convert the `happy`, `marital`, `degree`, and `health` columns into factors. Make sure the factor levels are arranged in a meaningful order.

2. Tabulate the status of (i) happiness, (ii) health, and (iii) happiness by health.

3. Create a barchart showing the *proportion* of happy respondents by marital status. The proportion is the ratio of the number of happy respondents over the number of respondents in the marital status group.

4. 
    1. Create a data frame containing respondents who are married and happy
    2. Create a data frame containing respondents who are either married or happy (or both)
    3. Find the number of respondents who are either married or happy, but not both
    4. Find the number of respondents who are neither married nor happy

5. Work with the `partyid` column as follows. 
    1. Set those who are `ind,near dem`, `not str democrat`, `strong democrat` to `democrat`
    2. Set those who are `ind,near rep`, `not str republican`, `strong republican` to `republican`
    3. Finally, tabulate the new `partyid` column